---
title: "p8105_hw3_ltc2121"
author: "Lindsey Covell"
output: github_document
---

```{r, echo = FALSE}
library(tidyverse)
library(p8105.datasets)
library(ggridges)
library(patchwork)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1 

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. Each row represents a single product from an instacart order.In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

There are a total of 134 aisles, with fresh vegetables and fresh fruits having the most items ordered from compared to all other aisles. 

```{r}
instacart %>% 
  count(aisle) %>% 
    arrange(desc(n))
```

Here is a plot which shows the number of items ordered in each aisle, starting with aisles that have at least 1000 items ordered from them. 

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) +
  geom_point() + 
    labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

This table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits` , and includes the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n)))  %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()

```

This table shows the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week in a format that is easy for viewing. 

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day, na.rm = TRUE)) %>%
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) %>% 
  knitr::kable(digits = 2)
```

## Problem 2 

```{r}
activity_df = 
  read_csv("hw3_data/accel_data.csv") %>% 
  pivot_longer(
    activity.1:activity.1440,
    names_to = "minute", 
    names_prefix = "activity.",
    values_to = "activity_count"
  ) %>% 
  mutate(
    activity_count = round(activity_count, 2), 
    day_type = ifelse(day == "Friday", "weekend", 
                     ifelse(day == "Saturday", "weekend", 
                            ifelse(day == "Sunday", "weekend", "weekday"))), 
  minute = as.numeric(minute), 
  day_type = as.factor(day_type))

```

This dataset contains `r nrow(activity_df)` rows and `r ncol(activity_df)` columns. Each row shows a distinct `activity count` for a given `minute` throughout a given `day` starting at midnight. The dataset spans over `r activity_df %>% select(week) %>% distinct %>% count` weeks and includes `r activity_df %>% select(day_id) %>% distinct %>% count` days of activity monitoring. The dataset also contains a variable that describes if the day constitutes as a weekend or a weekday. 

```{r}
activity_df %>% 
  group_by(week, day) %>% 
  summarise(total_activity = 
              sum(activity_count)) %>% 
  mutate(total_activity = round(total_activity, 0)) %>% 
  pivot_wider(names_from = day, values_from = total_activity ) %>% 
  knitr::kable()

######## arrange(desc(total_activity)) %>% ######## if using long data #### if not REMOVE !!
# 
```

The table above shows the total activity count for each day of the week. From this table you are able to see some general trends in activity. It seems like Saturday's are notably days with lower activity count, while Friday's seem to have higher activity counts. 

```{r}
activity_df %>% 
  ggplot(aes( x = minute, y = activity_count, color = day)) + 
  geom_point( alpha = .5) + 
  labs(
    title = "Activity during the Day by Day of the Week",
    x = "Hours of the Day",
    y = "Activity Count") + 
  scale_x_continuous(
    breaks = c(0, 360, 720, 1080, 1440), 
    labels = c("0", "6", "12", "18", "24"))
```

The graph above shows the activity count over the 24 hour day for each day in the dataset. The graph is grouped by day of the week. The graph shows that this individual typically starts their day around 7th-8th hour of the day. It is seen through outlying point that this person is more typical to have deviate from their average activity level on the weekend days. They typically have an increase in activity level between the 20th-22nd hour of the day, especially on Friday's. 

## Problem 3

```{r}
data("ny_noaa")
```

This dataset contains `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` columns. Each observation in the dataset is an individual daily reading from one of the New York state weather stations. The weather station readings were taken from `r min(ny_noaa$date)` to `r max(ny_noaa$date)`. The variables key measured from the weather station are precipitation, snowfall, snow depth, maximum temperature, and minimum temperature. There is a notable amount of missing data since not all of the weather stations measure each variable. In total, there are `r ny_noaa %>% mutate(tmax = as.numeric(tmax), tmin = as.numeric(tmin)) %>% pivot_longer(prcp:tmin, names_to = "type", values_to = "amount") %>% filter(is.na(amount)) %>% count()` missing values out of `r ny_noaa %>% mutate(tmax = as.numeric(tmax), tmin = as.numeric(tmin)) %>% pivot_longer(prcp:tmin, names_to = "type", values_to = "amount") %>% select(id, date, type) %>% distinct() %>% count()` values combined from all of the measured variables in the dataset. 

```{r}
ny_noaa %>% 
  mutate(year = lubridate::year(date), 
                month = lubridate::month(date), 
                day = lubridate::day(date), 
         tmax = as.numeric(tmax),
         tmin = as.numeric(tmin))

Hmisc::describe(ny_noaa$tmax)
```


