---
title: "p8105_hw3_ltc2121"
author: "Lindsey Covell"
output: github_document
---

```{r, echo = FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1 

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. Each row represents a single product from an instacart order.In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

There are a total of 134 aisles, with fresh vegetables and fresh fruits having the most items ordered from compared to all other aisles. 

```{r}
instacart %>% 
  count(aisle) %>% 
    arrange(desc(n))
```

Here is a plot which shows the number of items ordered in each aisle, starting with aisles that have at least 1000 items ordered from them. 

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) +
  geom_point() + 
    labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

This table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits` , and includes the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n)))  %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()

```

This table shows the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week in a format that is easy for viewing. 

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day, na.rm = TRUE)) %>%
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) %>% 
  knitr::kable(digits = 2)
```

